The Emergent Simplicity of Galaxy Size

### 2017-09-29 paper/galsize_paper.tex
Finished abstract. Too long

\begin{abstract}
We derive empirical modeling constraints on the connection between dark matter halos and the half-light radius $\rhalf$ of galaxies. Using forward-modeling techniques based on {\tt Halotools}, we confirm previous results in Kravtsov (2013) that $\rhalf$ is well-described by a linear scaling relation with halo virial radius. Novel to this work, we study galaxy size using new SDSS measurements of the $\rhalf-$dependence of galaxy clustering. Smaller galaxies cluster more strongly relative to larger galaxies of the same stellar mass, a new result. We show that the $\rhalf-$dependence of galaxy clustering is largely driven by the relative size of central and satellite galaxies: satellites are smaller than centrals of the same halo mass. We identify the ingredients that contribute most significantly to $\rhalf-$dependent clustering by building empirical models of galaxy size with various effects included and excluded. Stellar mass stripping of satellites has only a mild impact on $\rhalf-$dependent clustering and is unlikely to be the origin of the characteristic shape of the signal. On the other hand, correlations between galaxy size and host halo assembly history leave a dramatic imprint on the clustering signal that is strikingly similar in magnitude, $\mstar-$dependence and scale-dependence. In the class of models that give the most faithful recovery of the observed clustering signal, galaxy $\rhalf$ is set by the size of the  virial radius at the time the subhalo reaches its peak halo mass, with log-normal scatter in $\rhalf$ that is strongly correlated with the size of the halo scale radius at that time. Our results can be treated as a boundary condition for more complex and fine-grained models of galaxy size, and provide a simple means for cosmological surveys to generate synthetic galaxy populations with realistic sizes across the cosmic web.
\end{abstract}

>>>>

\begin{abstract}
We derive empirical modeling constraints on the connection between dark matter halos and the half-light radius of galaxies, $\rhalf.$ Novel to this work, we study galaxy size using new SDSS measurements of the $\rhalf-$dependence of galaxy clustering. We find that smaller galaxies cluster stronger relative to larger galaxies of the same stellar mass, a new result. We show that the $\rhalf-$dependence of galaxy clustering is largely driven by centrals being larger than satellite galaxies of the same halo mass. Stellar mass stripping of satellites has only a mild impact and is unlikely to be the origin of the characteristic magnitude and shape of the clustering signal. Much more successful models are based on the assumption that the mean galaxy profile {\em and its scatter} are correlated with the formation history of the dark matter halo profile. Based on our present findings, we predict that at fixed $\mhalo,$ galaxies with relatively shallow surface brightness profiles reside in relative low-concentration halos, and conversely. In the class of models that give the most faithful recovery of the observed clustering signal, galaxy $\rhalf$ is set by the size of the  virial radius at the time the subhalo reaches its peak halo mass, with log-normal scatter in $\rhalf$ that is strongly correlated with the halo scale radius $R_{\rm s}$ at that time. Our results can be treated as a boundary condition for more complex and fine-grained models of galaxy size, and provide a simple means for cosmological surveys to generate synthetic galaxy populations with realistic sizes across the cosmic web.
\end{abstract}
%Using forward-modeling techniques based on {\tt Halotools}, we confirm previous results in Kravtsov (2013) that $\rhalf$ is well-described by a linear scaling relation with halo virial radius.
